### 2013-11-12 Paper/paper.tex
Almost finished paper

\documentclass{llncs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{graphicx,color}
\usepackage{multicol}
\usepackage{bussproofs}
\usepackage{float}
\usepackage{centernot}
\usepackage{marginnote}

\usetikzlibrary{arrows,shapes,calc}

\DeclareFontFamily{U}{wncy}{}
\DeclareFontShape{U}{wncy}{m}{n}{<->wncyss10}{}
\DeclareSymbolFont{mcy}{U}{wncy}{m}{n}
\DeclareMathSymbol{\sh}{\mathord}{mcy}{"78}
\DeclareMathOperator{\sha}{\sh}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\fin}{finite}
\DeclareMathOperator{\rely}{rely}
\DeclareMathOperator{\guar}{guar}
\DeclareMathOperator{\letters}{letters}

\newcommand{\wlp}{\mathsf{wlp}}

\newcommand{\test}{\mathsf{test}}
\newcommand{\eval}{\mathsf{eval}}
\newcommand{\Id}{\mathsf{Id}}
\newcommand{\edn}{\mathsf{end}}
\newcommand{\unchanged}{\mathsf{unchanged}}
\newcommand{\preserves}{\mathsf{preserves}}

\newcommand{\udagger}{\text{$\rotatebox[origin=c]{180}{$\dagger$}$}}

\newtheorem{lem}{Lemma}

\begin{document}

\tikzstyle{elem} = [circle]
\tikzstyle{line} = [draw,thick, -latex']
\tikzstyle{rel} = [draw,thin,dashed, -latex']

\title{Algebraic Principles for Rely-Guarantee Style Concurrency Verification Tools}

\author{Alasdair Armstrong \and Victor B.~F.~Gomes \and Georg Struth}

\institute{Department of Computer Science, University of Sheffield, UK\\
\email{$\{$a.armstrong,v.gomes,g.struth$\}$@dcs.shef.ac.uk}}

\maketitle

\begin{abstract}
  We provide simple equational principles for deriving
  rely-guarantee-style inference rules and refinement laws based on
  idempotent semirings. We link the algebraic layer with concrete
  models of programs based on languages and execution traces. We have
  implemented the approach in Isabelle/HOL as a lightweight
  concurrency verification tool that supports reasoning about the
  control and data flow of concurrent programs with shared variables
  at different levels of abstraction. This is illustrated on two
  simple verification examples.
\end{abstract}

\pagestyle{plain}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

We aim to develop lightweight formal methods which support
rely-guarantee style reasoning about concurrent programs at varying
levels of abstraction. At the most abstract level, we use algebras as
a framework for reasoning about the control flow of programs as well
as for deriving inference rules and refinement laws. At the lowest
level, concrete models of program stores support fine-grained
reasoning about program data flow and inteference. We have implemented
this approach is the Isabelle/HOL interactive theorem proving
environment. Isabelle not only provides us with a high degree of
confidence in the correctness of our development and tool prototype,
it also supports the construction of custom proof tactics and
procedures for program verification and refinement tasks. Our approach
differs from existing frameworks for rely-guarantee reasoning in
Isabelle~\cite{nieto_rely-guarantee_2003}, both by making the link
between concrete programs and algebras explicit, and by allowing
arbitrary nested parallelism.

For sequential programs, the applicability of algebra, and Kleene
algebra in particular, has been known for over a decade. Kleene
algebra provides operations for non-deterministic choice, sequential
composition and finite iteration, in addition to skip and abort. With
appropriate extensions, Kleene algebras support Hoare-style program
verification, and allow the derivation of program equivalences and
refinement rules~\cite{kozen_kleene_1997,hoare_concurrent_2011}. Kleene algebras have
been applied in applications including compiler
optimisation~\cite{kozen_certification_2000}, program
construction~\cite{berghammer_automated_2010}, transformation and
termination analysis~\cite{desharnais_algebraic_2011}, and static
analysis~\cite{fernandes_describing_2007}. Formalisations and tools are
available in interactive theorem provers such as Coq~\cite{pous_kleene_2013} and
Isabelle~\cite{armstrong_kleene_2013,armstrong_program_2013,armstrong_algebras_2014}.

Algebra also plays a role in the study of separation logic, for
example, as the separation algebras in~\cite{calcagno_local_2007}, and in algebraic
separation logic~\cite{dang_algebraic_2011}. More recently,
concurrent Kleene algebras have given an algebraic account of some
aspects of concurrent separation logic~\cite{hoare_concurrent_2011,hoare_locality_2011} and
the first steps towards an algebraic description of rely-guarantee
based reasoning have been undertaken~\cite{hoare_concurrent_2011}.

One main characteristic of the rely-guarantee method is
compositionality. The verification of large concurrent programs is
reduced to the independent verification of individual subprograms,
where the effect of interferences from other subprograms are captured
by \emph{rely} and \emph{guarantee} conditions, which constrain the
behaviour of subprograms. This form of compositionality is, however,
weaker than the strong notion of compositionality inherent in algebra,
where full congruence rules exist for every operator and refinement is
usually based on monotonicity.

In this paper, we investigate algebraic principles for rely-guarantee
style reasoning. First step have already been taken, for example,
in~\cite{hoare_concurrent_2011}, but we extend this approach by extracting a minimal set
of axioms for relies and guarantees sufficient to derive the standard
rely-guarantee inference rules. Furthermore, we explore how the
algebraic compositionality of this approach can be broken in just the
right way, so as to capture the ideas of abstract interference
conditions in rely-guarantee reasoning.

The verification and construction of actual concurrent programs often
requires complex semantics involving synchronisation, termination
analysis and intricate data structures. For simplicity, in this paper
we eschew such models in favor of relatively simple language and trace
based semantics. Despite the simplicity of this model, we demonstrate
and evaluate our prototypical verification tool implemented in
Isabelle by verifiying simple two examples from the literature. Beyond
that our approach provides a coherent framework from which more
complex and detailed models can be integrated in the future.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algebraic Preliminaries}
\label{sec:KA}

Rely-guarantee algebras, which are introduced in the following
section, are based on dioids and Kleene algebras. A \emph{semiring} is
a structure $(S,+,\cdot,0,1)$ such that $(S,+,0)$ is a commutative
monoid, $(S,\cdot, 1)$ is a monoid and the distributivity laws $x\cdot
(y+z)=x\cdot z + y \cdot z$ and $(x+y)\cdot z = x\cdot z+y\cdot z$ as
well as the annihilation laws $x\cdot 0=0$ and $0\cdot x=0$ hold. A
\emph{dioid} is a semiring in which addition is idempotent:
$x+x=x$. Hence $(S,+,0)$ forms a join semilattice with least element
$0$ and partial order defined, as usual, as $x\le y\Leftrightarrow
x+y=y$. The operations of addition and multiplication are isotone with
respect to the order, that is, $x \le y $ implies $z+x\le z+y$,
$z\cdot x \le z\cdot y$ and $x\cdot z \le y\cdot z$. A dioid is
\emph{commutative} if multiplication is: $x\cdot y = y \cdot x$.

In the context of sequential programs, one typically thinks of $\cdot$
as sequential composition, $+$ as nondeterministic choice, $0$ as the
abortive action and $1$ as skip. In this context it is essential that
multiplication is not commutative. Often we use $;$ in place of
$\cdot$ when discussing programs. More formally, it is well known that
(regular) languages with language union as $+$ and language product as
$\cdot$, the empty language as $0$ and the empty word language
$\{\varepsilon\}$ as $1$ form dioids. Another model is formed by
binary relations with the union of relations as $+$, the product of
relations as $\cdot$, the empty relation as $0$ and the identity
relation as $1$. A model of commutative dioids is formed by sets of
(finite) multisets or Parikh vectors with multiset addition as
multiplication.

It is well known that commutative dioids can be used for modelling the
interaction between concurrent composition and nondeterministic
choice. The following definition serves as a basis for models of
concurrency in which sequential and concurrent composition
interact.

A \emph{trioid} is a structure $(S,+,\cdot,||,0,1)$ such that
$(S,+,\cdot,0,1)$ is a dioid and $(S,\cdot,||,0,1)$ a commutative
dioid. In a trioid there is no interaction between the sequential
composition $\cdot$ and the parallel composition $||$. On the one
hand, Gischer has shown that trioids are sound and complete for the
equational theory of series-parallel pomset languages~\cite{gischer_equational_1988},
which form a well studied model of true concurrency. On the other
hand, he has also obtained a completeness result with respect to a
notion of pomset subsumption for trioids with the additional
\emph{interchange axiom} $(w \| x) \cdot (y \| z) \le (w \cdot y) \|
(x \cdot z)$ and it is well known that this additional axiom also
holds for (regular) languages in which $||$ is interpreted as the
shuffle or interleaving operation~\cite{gischer_shuffle_1981}.

Formally, the \emph{shuffle} $\|$ of two finite words is defined
inductively as
\begin{equation*}
\epsilon \| s = \{s\}, \qquad s \| \epsilon = \{s\},\qquad
as \| bt = a(s \| bt) \cup b(as \| t),
\end{equation*}
which is then lifted to the shuffle product of languages $X$ and $Y$
as
\begin{equation*}
X \| Y = \{x \| y:\; x \in X \land x \in Y\}.
\end{equation*}

For logics of programs, notions of iteration or recursion are
essential. A \emph{Kleene algebra} is a dioid expaned with a star
operation which satisfies the \emph{left unfold axiom} $1 + x\cdot
x^*\le x^*$ and the \emph{left} and \emph{right induction axiom} $z+x
\cdot y \le y\Rightarrow x^*\cdot z \le y$ and $z + y\cdot x \le y
z\cdot \Rightarrow x^* \le y$. It follows that $1+x\cdot x^* = x^*$
and that the right unfold axiom $1+x^*\cdot x=x^*$ is derivable as
well. Thus iteration $x^\ast$ is modelled as the least fixpoint of the
function $\lambda y. 1+x\cdot y$, which is the same as the least
fixpoint of $\lambda y.1+ y\cdot x$. A \emph{commutative Kleene
  algebra} is a Kleene algebra in which multiplication is commutative.

It is well known that (regular) languages form Kleene algebras and
that (regular) sets of multisets form commutative Kleene algebras. In
fact, Kleene algebras are complete with respect to the equational
theory of regular languages as well as the equational theory of binary
relations with the reflexive transitive closure operation as the
star~\cite{kozen_completeness_1994}. Moreover, commutative Kleene
algebras are complete with respect to the equational theory of regular
languages over multisets~\cite{conway_regular_1971}. It follows that
equations in (commutative) Kleene algebras are decidable.

A \emph{bi-Kleene algebra} is a structure
$(K,+,\cdot,||,0,1,\hphantom{}^*,\hphantom{}^{(*)})$ where
$(K,+,\cdot,0,1,\hphantom{}^*)$ is a Kleene algebra and
$(K,+,||,0,1,\hphantom{}^{(*)})$ is a commutative Kleene
algebra. Bi-Kleene algebras are sound and complete with respect to the
equational theory of regular series-parallel pomset languages, and the
equational theory is again decidable~\cite{LaurenceStruth}. A
\emph{concurrent Kleene algebra} is a bi-Kleene algebra which
satisfies the interchange law~\cite{hoare_concurrent_2011}. It can be shown that
shuffle languages and regular series-parallel pomset languages with a
suitable notion of pomset subsumption form concurrent Kleene algebras.

In some contexts, it is also useful to add a meet operation $\sqcap$,
such that $(D,+,\sqcap)$ is a distributive lattice. This is
particularly needed in the context of refinement, where we typically
want to represent specifications as well as programs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Generalised Hoare Logics  in Kleene Algebra}

It is well known that the inference rules for sequential Hoare logic
(except the assignment axiom) can be derived in expansions of Kleene
algebras. One approach is as follows~\cite{moller_algebras_2006}. Suppose a
suitable Boolean algebra $B$ of \emph{tests} has been embedded into a
Kleene algebra $K$ such that $0$ and $1$ are the minimal and maximal
element of $B$, $+$ corresponds to join and $\cdot$ to
meet. Complements $-$ are defined only on $B$. Suppose further that a
\emph{backward diamond operator} $\langle x|p$ has been defined for
each $x\in K$ and $p\in B$, which models the set of all states to
which each terminating execution of program $x$ may lead from states
$q$. Finally suppose that a \emph{forward box operator} $|x]p$ has
been defined which models the (largest) set of states from which every
terminating execution of $x$ must end up in states $q$ and that boxes
and diamonds are adjoints of the Galois connection
\begin{equation*}
  \langle x|p \le q \Leftrightarrow p \le |x]q,
\end{equation*}
for all $x\in K$ and $p,q\in B$. It is then evident from the above
explanations that validity of a Hoare triple $\vdash \{p\}x\{q\}$ can
be encoded as $\langle x|p \le q$ and the weakest liberal precodition
operator $\mathsf{wlp}(x,q)$ as $|x]p$. Hence the relationship between
the proof theory and the semantics of Hoare logic is captured by the
Galois connection $ \vdash\{p\}x\{q\} \Leftrightarrow p \le
\mathsf{wlp}(x,q)$. It has been shown that the relational semantics of
sequential while-programs can be encoded in these \emph{modal Kleene
  algebras} and that the inference rules of Hoare logic can be
derived~\cite{moller_algebras_2006}.

In the context of concurrency, this relational approach is no longer
appropriate. Here, the following approach by
Tarlecki~\cite{tarlecki_language_1985} can be used instead. One can
now encode validity of a Hoare triple as
\begin{equation*}
  \vdash \{x\}y\{z\} \Leftrightarrow x \cdot y \le z
\end{equation*}
for arbitrary elements of a Kleene algebra $K$. Nevertheless all the
rules of sequential Hoare logic except the assignment axiom can still
be derived in this setting~\cite{hoare_concurrent_2011}. Tarlecki's motivating
explanations carry over to the algebraic approach.

As an example we show the derivation of a generalised while
rule. Suppose $x\cdot t\cdot y\le x$. Then $x\cdot (t\cdot y)^\ast \le
x$ by the right induction axiom of Kleene algebra and therefore
$x\cdot (t\cdot y)^\ast t'\le x\cdot t'$ for arbitrary element $t'$ by
isotonicity of multiplication. This derives the while rule
\begin{equation*}
  \frac{\vdash \{x\cdot t\}y\{x\}}{\vdash \{x\}(t\cdot y)^\ast\{t'\cdot x\}}
\end{equation*}
for a generalised while loop $(t\cdot y)^\ast\cdot t'$, which
specialises to the conventional rule when $t$ and $t'$ are, in some
sense, complements.

The correspondence to a wlp-style semantics, as in model Kleene
algebra, now requires a generalisation of the Galois connection for
boxes and diamonds to multiplication and an upper adjoint in the form
of residuation.  This can be achieved in the context of \emph{action
  algebras}~\cite{pratt_action_1990}, which expand Kleene algebras by operations
of left and right residuation defined by the Galois connections
\begin{equation*}
  x\cdot y\le z\Leftrightarrow x\le z\leftarrow y,\qquad x\cdot y\le z \Leftrightarrow y\le x\rightarrow z.
\end{equation*}
These residuals, and now even the Kleene star can be axiomatised
equationally in action algebras. For a comprehensive list of the
properties of action algebras and their most important models
see~\cite{armstrong_kleene_2013}, including the language and the relational
model. In analogy to the development in model Kleene algebra we can
now stipulate $ \wlp(x,y)=y\leftarrow x$ and obtain the Galois
connection
\begin{equation*}
  \vdash \{x\}y\{z\} \Leftrightarrow x\le \wlp(y,z)
\end{equation*}
with the characteristic properties $\vdash \{\wlp(y,z)\}y\{z\}$ and $x\le
\wlp(y,z)\Rightarrow \ \vdash \{x\}y\{z\}$. Moreover, if the action algebra is
also a quantale, and infinite sums exist, it follows that $
\wlp(y,z)=\sum\{x:\ \vdash \{x\}y\{z\}\}$.  It is obvious that this definition
makes sense in all models of action algebras and
quantales. Intuitively, suppose $p$ stands for the set of all
behaviours of a system, for instance the set of all execution traces,
that end in state $p$, and likewise for $q$. Then $\{p\}x\{q\}$ states
that all executions ending in $p$ can be extended by $x$ to executions
ending in $q$. $\wlp(x,q)$ is the most general behaviour, that is the
set of all executions $p$ after which all executions of $x$ must end
in $q$.

In the context of concurrent Kleene algebra, even without the parallel
star, additional inference rules can be derived, in particular
\begin{equation*}
  \frac{\vdash\{x\}y\{z\}}{\vdash\{w||x\}y\{w||z\}}\qquad \frac{\vdash\{x_1\}y_1\{z_1\}\quad\vdash\{x_2\}y_2\{z_2\}}{\vdash\{x_1||x_2\}y_1||y_2\{z_1||z_2\}}
\end{equation*}
which correspond to the frame rule and the concurrency rule of
concurrent separation logic~\cite{} when $||$ is interpreted as
separating conjunction. In fact, the second inference rule is, in any
bi-Kleene algebra equivalent to the interchange axiom and the frame
rule is derivable from this axiom as well~\cite{hoare_concurrent_2011}. The
precise connection between concurrent separation logic, a variant of
concurrent Kleene algebra and a resource transformer model has been
elaborated in~\cite{hoare_locality_2011}.

A residuation for concurrent composition can be considered as well:
\begin{equation*}
x\|y \le z \Leftrightarrow y \le x/z.
\end{equation*}
The residual $x/z$ represents the weakest program such that when
placed in parallel with $x$, the parallel composition behaves as $z$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{A  Rely-Guarantee Algebra}
\label{sec:RG}

In this section we show how bi-Kleene algebras can be expanded into a
simple algebra that supports the derivation of rely-guarantee-style
inference rules. This development does \emph{not} use the interchange
law for several reasons. First, this law fails for fair parallel
composition $x\parallel_f y$ in a model with possibly infinite, or
non-terminating programs. In this model, $x \cdot y \not\leq x
\parallel_f y$ in the case whenever $x$ is non-terminating. Secondly,
it is not needed for deriving the usual laws or rely-guarantee.

A rely-guarantee algebra is a structure
$(K,C,+,\sqcap,\cdot,\|,^\star,0,1)$, where $(K,+,\sqcap)$ is a
distributive lattice, $(K,+,\cdot,\|,0,1)$ is a trioid and
$(K,+,\cdot,^\star,0,1)$ is a Kleene algebra. Note that we do not
consider the parallel star. $C$ is a distinguished subset of rely and
guarantee conditions or \emph{interference constraints} which satisfy
the following axioms
\begin{align}
r\|r &\le r, \label{rg1}\\
r &\le r\|r', \label{rg2}\\
r\|xy &= (r\|x)(r\|y), \label{rg3}\\
r\|x^+ &\le (r\|x)^+ \label{rg4}.
\end{align}
By convention, we use $r$ and $g$ to refer to elements of $C$,
depending on whether they are used as relies or guarantees, and
$x,y,z$ for arbitrary elements of $K$. The operators $\|$ and $\sqcap$
must be closed with respect to $C$.

Axiom (\ref{rg1}) states that interference from a rely condition being
run twice in parallel is no different from just the interference from
that rely condition begin run once in parallel. Axiom (\ref{rg2})
states that interference from a single rely condition is less than
interference from itself and another rely condition. Axiom (\ref{rg3})
allows a rely or guarantee to be split across sequential
programs. Axiom (\ref{rg4}) is similar to Axiom (\ref{rg3}) in intent,
except it deals with finite iteration.

Some elementary consequences of these rules are
\begin{equation*}
1 \le r,\qquad
r^\star = rr = r = r\|r,\qquad
r\|x^+ = (r\|x)^+.
\end{equation*}

Axioms (\ref{rg1}), (\ref{rg2}) and (\ref{rg3}) are independent. To
show this, we use Isabelle's
\emph{Nitpick}~\cite{blanchette_nitpick:_2010} counterexample
generator to construct models which violate each particular axiom
while satisfying all others.

Axiom (\ref{rg4}) can be derived from (\ref{rg3}) by a simple
inductive argument in quantales where the plus operation can be
defined as a sum of powers $x^+=\sum_{i\ge 1} x^i$ where $x^1=x$
and $x^{i+1}=x\cdot x^i$. In first-order Kleene algebras, however, it
is independent, but it is impossible to find a counterexample with Nitpick because all such counterexamples are finite, but all finite Kleene algebras are a forteriori quantales.

Jones quintuples can be encodeded in this setting as
\begin{align}
r, g \vdash \{p\} c \{q\} \iff p(r\|c) \le q \land c \le g. \label{quin}
\end{align}
With this encoding we can derive the standard rely-guarantee inference
rules, as shown in Figure \ref{fig:rgrules}. Thus equations
(\ref{rg1}) to (\ref{rg4}), which are all necessary to derive these
rules, somehow represent a minimal set of axioms from which these
inference rules can be derived.

By using these residuals quintuples can be encoded in the following way,
which is equivalent to the encoding in Equation (\ref{quin}).
\begin{align}
r, g \vdash \{p\} c \{q\} \iff c \le r/(p \rightarrow q) \sqcap g \label{refine}.
\end{align}
This encoding allows us to think in terms of program refinement~\cite{hayes_refining_2013}, as
$r/(p \rightarrow q) \sqcap g$ defines the weakest program that when
placed in parallel with interference from $r$, and guaranteeing
interference at most $g$, goes from $p$ to $q$---a generic
specification for a concurrent program.

\begin{figure}[tbh]
\centering
\begin{prooftree}
\RightLabel{Skip}
\AxiomC{$pr \le p$}
\UnaryInfC{$r, g \vdash \{p\}1\{p\}$}
\end{prooftree}

\begin{prooftree}
\RightLabel{Weakening}
\AxiomC{$r' \le r$}
\AxiomC{$g \le g'$}
\AxiomC{$p \le p'$}
\AxiomC{$r', g' \vdash \{p'\}x\{q'\}$}
\AxiomC{$q' \le q$}
\QuinaryInfC{$r, g \vdash \{p\}x\{q\}$}
\end{prooftree}

\begin{prooftree}
\RightLabel{Sequential}
\AxiomC{$r, g \vdash \{p\}x\{q\}$}
\AxiomC{$r, g \vdash \{q\}y\{s\}$}
\BinaryInfC{$r, g \vdash \{q\}xy\{s\}$}
\end{prooftree}

\begin{prooftree}
\RightLabel{Parallel}
\AxiomC{$r_1, g_2 \vdash \{p_1\}x\{q_1\}$}
\AxiomC{$g_1 \le r_2$}
\AxiomC{$r_1, g_2 \vdash \{p_2\}y\{q_2\}$}
\AxiomC{$g_2 \le r_1$}
\QuaternaryInfC{$r_1 \sqcap r_2, g_1 \| g_2 \vdash \{p_1 \sqcap q_2\}x\|y\{q_1 \sqcap q_2\}$}
\end{prooftree}

\begin{prooftree}
\RightLabel{Choice}
\AxiomC{$r, g \vdash \{p\}x\{q\}$}
\AxiomC{$r, g \vdash \{p\}y\{q\}$}
\BinaryInfC{$r, g \vdash \{p\}x + y\{q\}$}
\end{prooftree}

\begin{prooftree}
\RightLabel{Star}
\AxiomC{$pr \le p$}
\AxiomC{$r, g \vdash \{p\}x\{p\}$}
\BinaryInfC{$r, g \vdash \{p\}x^\star\{p\}$}
\end{prooftree}
\caption{Rely-guarantee inference rules}
\label{fig:rgrules}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Breaking Compositionality}
\label{sec:INT}

While the algebra in section \ref{sec:RG} is adequate for deriving the
standard inference rules, it's equality is too strong to capture many
interesting statements about concurrent programs. Consider the
congruence rule for parallel composition, which is inherent in the
algebraic approach:
\begin{align*}
x = y \implies x\|z = y\|z.
\end{align*}
This can be read as follows; if $x$ and $y$ are equal, then they must
be equal under all possible interference from any arbitrary $z$. At
first sight, this might seem to preclude any reasoning about
interference in a fine-grained way using purely algebra. This is not
the case, but breaking such inherent compositionality in just the
right way to capture interesting properties of concurrency requires
extra work.

The most obvious way we can acheive this is by expanding our
rely-guarantee algebra with an additional function $\pi$ and redefining
our quintuples as,
\begin{align*}
r, g \vdash \{p\} c \{q\} \iff \pi(p(r\|c)) \le \pi(q) \land c \le g.
\end{align*}
Since for any operator $\bullet$, it is not required that
\begin{align*}
\pi(x) = \pi(y) \implies \pi(x \bullet z) = \pi(y \bullet z),
\end{align*}
we can `break' compositionality in just the right way, provided we
choose the correct properties for $\pi$. For example, if we consider
the following seven axioms
\begin{align}
\pi(\pi(x)) &= \pi(x), \label{pi1}\\
x \le y \implies \pi(x) &\le \pi(y), \label{pi2}\\
\pi(x) &\le x, \label{pi3}\\
\pi(x^*) &\le \pi(\pi(x)^*), \label{pi4}\\
\pi(xy) &\le \pi(\pi(x)\pi(y)), \label{pi5}\\
\pi(x \sqcap y) &= \pi(x) \sqcap \pi(y). \label{pi6}
\end{align}

Axioms (\ref{pi1}) to (\ref{pi3}) show that $\pi$ is an interior
operator, while (\ref{pi4}) and (\ref{pi5}) allow us to introduce
$\pi$ as needed around sequential composition and finite iteration,
which allows us to substitute under those operators as required---we
only want to break compositionality for $\|$.

The above axioms indicate that we could define $\pi$ as a function
$\lambda x.\; x \sqcap c$ where $c$ satisfies
\begin{align}
x^* \sqcap c &\le (x \sqcap c)^* \sqcap c, \label{con1}\\
xy \sqcap c &\le (x \sqcap c)(y \sqcap c) \sqcap c. \label{con2}
\end{align}

As such we redefine our rely-guarantee algebra from section
\ref{sec:RG} as a structure $(K,RG,+,\sqcap,\cdot,\|,^\star,0,1,c)$
which, in addition to the rules in section \ref{sec:RG} satisfies both
(\ref{con1}) and (\ref{con2}).

All the rules in Figure \ref{fig:rgrules} can be derived with this
algebra, and moreover their proofs remain the same, mutatis
mutandis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Language Model}
\label{sec:Model}

In this section, we construct a finite language model satisfying the
axioms in Section \ref{sec:RG}. Restricting our attention to finite
languages means we do not need to concern ourselves with non-algebraic
termination side-conditions, nor do we need to worry about additional
restrictions on parallel composition, e.g. fairness. However, all the
results in this section can easily be adapted to potentially
infinite languages, and our Isabelle/HOL formalisation includes these more
general definitions by using coinductively defined lazy lists to
represent words, and having a weakly-fair shuffle operator for such
infinite languages.

We consider languages where the alphabet contains state pairs of the
form $(\sigma_1,\sigma_2) \in \Sigma \times \Sigma$. A word in such a
language is \emph{consistent} if every such pair in a word has the
same first state as the previous transition's second state. For
example, $(\sigma_1,\sigma_2)(\sigma_2,\sigma_3)$ is consistent, while
$(\sigma_1,\sigma_2)(\sigma_3,\sigma_3)$ is inconsistent. Sets of
consistent words are essentially \emph{Aczel
  traces}~\cite{boer_formal_1999}, but lack the usual process
labels. We denote the set of all consistent words by $T$.

Sequential composition in this model is language product, as per
usual. Concurrent composition is the shuffle product defined in
section \ref{sec:KA}. The shuffle product is associative, commutative,
and distributes over arbitrary joins. Both products share the same
unit, $\{\epsilon\}$ and zero, $\emptyset$. Proving properties of
shuffle is suprisingly tricky (especially if one considers infinite
words). For a in-depth treatment of the shuffle product
see~\cite{mateescu_shuffle-like_1997}.

\begin{lem}
$(\mathcal{P}((\Sigma\times\Sigma)^*), \cup, \cdot,\sha, \emptyset,
  \{\epsilon\})$ forms a trioid.
\end{lem}

The rely-guarantee elements in this model are sets containing all the
words which can be built from some set of state pairs in
$\Sigma\times\Sigma$. We define a function $\langle R\rangle$ which
lifts a relation $R$ to a language containing words of length one for
each pair in $R$. The set of rely-guaratee conditions $C$ is then defined
as $\{r.\; \exists R. r = \langle R\rangle^*\}$.

\begin{lem}
$(\mathcal{P}((\Sigma\times\Sigma)^*), C, \cup, \cdot,\sha,
  \emptyset, \{\epsilon\}, T)$ is a rely-guarantee algebra.
\end{lem}
Since $\langle R\rangle$ is atomic, it satisfies several useful properties, such as,
\begin{align*}
\langle R\rangle^\star \| \langle S\rangle &= \langle R\rangle^\star; \langle S\rangle; \langle R\rangle^\star\\
\langle R\rangle^\star \| \langle S\rangle^\star &= (\langle R\rangle^\star; \langle S\rangle^\star)^\star
\end{align*}

To demonstrate how this model works, consider the graphical
representation of a language shown below.

\begin{center}
\begin{tikzpicture}[x=1cm,auto]
  \node (center) {};
  \node [elem] (s12) {$\sigma_2$};
  \node [elem, above of=s12] (s11) {$\sigma_1$};
  \node [elem, below of=s12] (s13) {$\sigma_3$};
  \node [right of=s13, node distance=0.75cm] (r1) {};

  \node [elem, right of=s12, node distance=1.5cm] (s22) {$\sigma_2$};
  \node [elem, above of=s22] (s21) {$\sigma_1$};
  \node [elem, below of=s22] (s23) {$\sigma_3$};
  \node [right of=s23, node distance=0.75cm] (r2) {};

  \node [elem, right of=s22, node distance=1.5cm] (s32) {$\sigma_2$};
  \node [elem, above of=s32] (s31) {$\sigma_1$};
  \node [elem, below of=s32] (s33) {$\sigma_3$};
  \node [right of=s33, node distance=0.75cm] (r3) {};

  \node [elem, right of=s32, node distance=1.5cm] (s42) {$\sigma_2$};
  \node [elem, above of=s42] (s41) {$\sigma_1$};
  \node [elem, below of=s42] (s43) {$\sigma_3$};
  \node [right of=s43, node distance=0.75cm] (r4) {};

  \path [line] (s11) -- (s21);
  \path [line] (s21) -- (s32);
  \path [line] (s32) -- (s43);

  \path [rel] (s11) -- (s22);
  \path [rel] (s12) -- (s22);
  \path [rel] (s23) -- (s32);
\end{tikzpicture}
\end{center}
The language contains the following six words
\begin{align*}
(\sigma_1,\sigma_1)(\sigma_1,\sigma_2)(\sigma_2,\sigma_3), \qquad
(\sigma_1,\sigma_2)(\sigma_1,\sigma_2)(\sigma_2,\sigma_3),\\
(\sigma_2,\sigma_2)(\sigma_1,\sigma_2)(\sigma_2,\sigma_3), \qquad
(\sigma_1,\sigma_1)(\sigma_3,\sigma_2)(\sigma_2,\sigma_3),\\
(\sigma_1,\sigma_2)(\sigma_3,\sigma_2)(\sigma_2,\sigma_3), \qquad
(\sigma_2,\sigma_2)(\sigma_3,\sigma_2)(\sigma_2,\sigma_3),
\end{align*}
where only the first,
$(\sigma_1,\sigma_1)(\sigma_1,\sigma_2)(\sigma_2,\sigma_3)$ is
consistent. This word is highlighted with solid arrows in the
diagram above. Now if we shuffle the single state pair $(\sigma_2,
\sigma_3)$ into the above language, we might end up with a language as
below:

\begin{center}
\begin{tikzpicture}[x=1cm,auto]
  \node [elem] (s12) {$\sigma_2$};
  \node [elem, above of=s12] (s11) {$\sigma_1$};
  \node [elem, below of=s12] (s13) {$\sigma_3$};
  \node [right of=s13, node distance=0.75cm] (r1) {};

  \node [elem, right of=s12, node distance=1.5cm] (i2) {$\sigma_2$};
  \node [elem, above of=i2] (i1) {$\sigma_1$};
  \node [elem, below of=i2] (i3) {$\sigma_3$};
  \node [right of=i3, node distance=0.75cm] (ri) {};

  \node [elem, right of=i2, node distance=1.5cm] (s22) {$\sigma_2$};
  \node [elem, above of=s22] (s21) {$\sigma_1$};
  \node [elem, below of=s22] (s23) {$\sigma_3$};
  \node [right of=s23, node distance=0.75cm] (r2) {};

  \node [elem, right of=s22, node distance=1.5cm] (s32) {$\sigma_2$};
  \node [elem, above of=s32] (s31) {$\sigma_1$};
  \node [elem, below of=s32] (s33) {$\sigma_3$};
  \node [right of=s33, node distance=0.75cm] (r3) {};

  \node [elem, right of=s32, node distance=1.5cm] (s42) {$\sigma_2$};
  \node [elem, above of=s42] (s41) {$\sigma_1$};
  \node [elem, below of=s42] (s43) {$\sigma_3$};
  \node [right of=s43, node distance=0.75cm] (r4) {};

  \path [rel] (s11) -- (i1);
  \path [rel] (s21) -- (s32);
  \path [line] (s32) -- (s43);
  \path [line] (i2) -- (s23);

  \path [line] (s11) -- (i2);
  \path [line] (s12) -- (i2);
  \path [line] (s23) -- (s32);
\end{tikzpicture}
\end{center}

By performing this shuffle action, we no longer have a consistent word
from $\sigma_1$ to $\sigma_3$, but instead a consistent word from
$\sigma_2$ to $\sigma_3$ and $\sigma_1$ to $\sigma_3$. These new
consistent words were constructed from previously inconsistent
words---the shuffle operator can take two inconsistent words and
shuffle them together in such a way to generate many consistent
words. If we only considered consistent words, \'a la Aczel traces, we
would be unable to define such a shuffle operator directly on the
traces themselves, and would instead have to rely on some operational
semantics to generate traces.

Following~\cite{brookes_full_1993} and~\cite{dingel_refinement_2002},
we inductively generate the mumble language $w^\dagger$ for a word $w$
in a language over $\Sigma^2$ as follows: Assume $R,S \in \Sigma^2$
and $u,v,w \in (\Sigma^2)^*$. First, $w \in w^\dagger$. Secondly, if
$uRSv \in w^\dagger$ then $u(R;S)v \in w^\dagger$. This operation is
lifted to languages in the obvious way as
\begin{align*}
X^\dagger = \bigcup\{x^\dagger.\; x \in X\}.
\end{align*}
Stuttering is represented as a rely condition $\langle \Id\rangle^\star$
where $\Id$ is the identity relation. Two languages $X$ and $Y$ are
equal under stuttering if $\langle \Id\rangle^\star \| X =_\pi \langle
\Id\rangle^\star \| Y$.

Stuttering and mumbling are important for defining \emph{tests}. A
test is any language $P$ where $P \le \langle \Id\rangle$. We
write $\test(P)$ for $\langle \Id_P\rangle$. Without any
stuttering and mumbling the traces $\test(P); \test(Q)$ and
$\test(P \cap Q)$ are incomparable, as all words in the former have
length two, while all the words in the latter have length one. With mumbling we have that
\begin{align*}
\test(P \cap Q) \le_\pi \test(P); \test(Q)
\end{align*}
as the longer words in $\test(P);\test(Q)$ can be mumbled down into
the shorter words of $\test(P\cap Q)$, whereas with stuttering we
have the opposite,
\begin{align*}
\langle I\rangle^\star \| (\test(P);\test(Q)) \le_\pi \langle I\rangle^\star \| \test(P \cap Q).
\end{align*}
For the rest of the paper, we assume that all languages are
implicitly mumble closed.

Using tests, we can encoding common program statements such as if
statements and while loops as in Kozen's Kleene algebra with tests.
\begin{align*}
\mathsf{if } P \text{ \{ } X \text { \} \textsf{else} \{ } Y \text{ \}} &= \test(P); X + \test(- P); Y\\
\mathsf{while } P \text{ \{ } X \text{ \}} &= (\test(P);X)^\star;\test(- P)
\end{align*}
Finally, we define the operator $\edn(P)$ which is contains all the
words which end in a state satisfying $P$. Some useful properties of $\edn$ include
\begin{align*}
\edn(P); \test{Q} &\le_\pi \edn(P \cap Q),\\
\test(P) &\le \edn(Q),\\
\text{range}(\Id_P \circ R) \le P \implies \edn(P); \langle R\rangle^\star &\le_\pi \edn(P).
\end{align*}

In the trace model, assignment is defined as
\begin{align*}
  x := e = \bigcup v.\,\test\{\sigma.\,\eval(\sigma,e) = v\} \cdot x \leftarrow v
\end{align*}
where $x \leftarrow v$ denotes the atomic command which assigns the
value $v$ to $x$. The $\eval$ function evaluates an expression $e$ in
the state $\sigma$. Using this definition of assignment we can derive
the assignment rule
\begin{align*}
&\unchanged(\mathsf{vars}(e)) \cap \preserves(P) \cap \preserves(P[x/e]),\\
&\unchanged(- \{x\})\\
&\vdash \{\edn(P)\} x := e \{\edn(P[x/e])\}.
\end{align*}

The rely condition states the following: First, the environment is not
allowed to modify any of the variables used when evaluating $e$,
i.e. those variables must remain unchanged. Second, that the
environment must preserve the precondition. Third, that postcondition
of the assignment statement is also preserved. In turn, the assignment
statement itself guarantees that it leaves every variable other than
$x$ unchanged. Preserves and unchanged are defined as
\begin{align*}
\preserves(P) &= \langle \{(\sigma,\sigma').\; P(\sigma) \implies P(\sigma')\}\rangle^\star,\\
\unchanged(X) &= \langle \{(\sigma,\sigma').\; \forall v\in X.\; \sigma(v) = \sigma'(v)\}\rangle^\star.
\end{align*}
We also defined two futher rely conditions, increasing and decreasing,
which are defined much like unchanged except they only require that
variables increase or decrease, rather than stay the same. Note that
we can easily define other useful assignment rules---if we know things
about $P$ and $e$, we can make stronger guarantees about what $x := e$
can do. For example the assignment $x := x - 2$ can also guarantee
that $x$ will always decrease.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}

\fbox{Simple assignment example}

Figure \ref{fig:findp} shows the classic FINDP program, which has been
used by numerous authors including Owicki~\cite{owicki_axiomatic_1975},
Jones~\cite{jones_development_1981}, de Roever~\cite{de_roever_concurrency_2001} and
Hayes~\cite{hayes_refining_2013}. The program finds the least element of an array
satisfying the predicate $P$. The index of the first element
satisfying $p$ is placed in the variable $f$. If there are no elements
of the array satisfying $P$ then $f$ will be set to the length of the
array. The program has two subprograms, $A$ and $B$, which run in
parallel, one of which searches the even indices while the other
searches the odd indices. A speedup over a sequential implementation
is acheived as $A$ will terminate when $B$ finds an element of the
array satisfying $P$ which is less than $i_A$.
\begin{figure}
\[
\begin{aligned}
&f_A := \mathsf{len}(\mathsf{array});\\
&f_B := \mathsf{len}(\mathsf{array});\\
&\left(
\begin{aligned}
&i_A = 0\\
&\text{\textsf{while} } i_A < f_A \land i_A < f_B \text{ \{ }\\
&\qquad\text{\textsf{if} } P(\mathsf{array}[i_A]) \text{ \{}\\
&\qquad\qquad f_A := i_A\\
&\qquad\text{\} \textsf{else} \{}\\
&\qquad\qquad i_A := i_A + 2\\
&\qquad\text{\}}\\
&\text{\}}
\end{aligned}
\middle\|
\begin{aligned}
&i_B = 1\\
&\text{\textsf{while} } i_B < f_A \land i_B < f_B \text{ \{}\\
&\qquad\text{\textsf{if} } P(\mathsf{array}[i_B]) \text{ \{}\\
&\qquad\qquad f_B := i_B\\
&\qquad\text{\} \textsf{else} \{}\\
&\qquad\qquad i_B := i_B + 2\\
&\qquad\text{\}}\\
&\text{\}}
\end{aligned}
\right);\\
&f = \text{min}(f_A,f_B)
\end{aligned}
\]
\caption{FINDP Program}
\label{fig:findp}
\end{figure}

Here, we sketch the main outline of the correctness proof, and comment
on its implementation in Isabelle. We don't attempt to give a fully
detailed proof, as this has been done many times previously.

To prove the correctness of FINDP, we must show that
\begin{align*}
\text{FINDP} \le_\pi \edn(\mathsf{leastP}(f)) + \edn(f = \mathsf{len}(\text{array})).
\end{align*}
where $\mathsf{leastP(f)}$ is the set of states where $f$ is the least
index satisfying $P$, and $f = len(array)$ is the set of states where
$f$ is the length of the array. In other words, either we find the
least element, or no elements in the array satisfy $P$, in which case
$f$ remains the same as the length of the array.

To prove the parallel part of the program, subprogram $A$ guarantees
that it does not modify any of the variables used by subprogram $B$,
except for $f_A$, which it guarantees will only ever
decrease. Subprogram $B$ makes effectively the same guarantee to
$A$. Under these interference constraints we then prove that $A$ or
$B$ will find the lowest even or odd index which satisfies $P$
respectively---or they do not find it, in which case $f_A$ or $f_B$
will remain equal to the length of the array.

Despite the seemingly straightforward nature of this proof, this turns
out to be suprising difficult to prove in Isabelle. Each atomic step
needs to be shown to satisfy the guarantee of its containing
subprogram, as well as any goals relating to it's pre and post
conditions. This invariably leads to a proliferation of many numerous
proof goals, even for such a small program. More work needs to be done
to manage the complexity of such proofs within interactive theorem
provers. Existing work on combining rely-guarantee with separation
logic~\cite{vafeiadis_modular_2008} may proof useful here. Our language model is
sufficiently generic such that arbitrary models of stores may be used,
including those common in separation logic, which have already been
implemented in Isabelle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

CNPq EPSRC

\bibliography{paper}{}
\bibliographystyle{plain}

\end{document}